Sparse Matrix Vector Multiplication with Multiple vectors and Multiple Matrices on the MIC Architecture

### 2013-10-14 ieee_paper.tex
temp

Registers on the MIC are 512 bits wide, and can store 16 floats or 8
doubles. It is through these registers that the \#SIMD pragma is able
to achieve vectorization. (In comparison to OpenCL or CUDA, one can
think of these registers as a warp.)  We use the following vector
operations in our code, which we'll explain in the text.

@@

\begin{figure*}
  %
  \begin{center}
    \begin{tabular}{|l|l|}
      \hline
      \loadps &  load 16 consecutive floats to register\\
      \fmadps &  multiply/add of 16 floats\\
      \gatherps &  gather 16, possibly disconnected floats from memory\\
      \swizzleps &  reorder a channel, and duplicate across channels\\
      \storenrngops &  store 16 floats to memory without register rewrite or reordering\\
      \castpstoi & reinterpret 16 floats as integers\\
      \castitops & reinterpret 16 integers as floats\\
      \permute &  permute channels; do not change individual channels\\
      \intmask &  \\
      \loadunpack &  \\
      \hline
    \end{tabular}
  \end{center}
  %
  \caption{List of MIC vectorial instructions used}
\end{figure*}
In the vector instructions, \ttt{ps} refers to a float (4 bytes),
while \ttt{epi32} refers to a 4-byte integer. Although we do not
discuss double precision in this paper, \ttt{ds} refers to an 8 byte
double precision real number (i.e., \loadds).  Each vector register is
broken up into four channels of 128 bits each. It is possible to
interchange these channels at a cost of a "few" \qes{exact numbers?}
cycles, and it is also possible to execute swizzle operations within a
channel. For example, if the 128 bytes of each channel are labelled as
$ABCD$, the vector instruction \ttt{\_mm512\_swizzle\_ps(v,
  \_MM\_SWIZ\_REG\_AAAA)} replaces each channel by $AAAA$.

>>>>

As explained in Section~XXX, we expect to achieve a performance of
about 210Gflop/s in single precision in the best case. We will focus
on the single precision case with 4 vectors and 4 matrices. But
similar techniques apply for other combinations. This target
performance represents 10\% of the peak performance of the
architecture which is impossible without an efficient vectorization of
the kernel. Indeed, that means a Fused Multiply-Add instruction on  
fully loaded registers must be executed at most every 10 cycles.

Registers on the MIC are 512 bits wide, and can store 16 floats or 8
doubles. It is through these registers that the \#SIMD pragma is able
to achieve vectorization. (In comparison to OpenCL or CUDA, one can
think of these registers as a warp.)  We use the vector operations
presented in Figure~\ref{fig:MIC-vect-instr} in our code, which we'll
explain in the text.

@@

\begin{figure*}
  %
  \begin{center}
    \begin{tabular}{|l|l|}
      \hline
      \loadps &  load 16 consecutive floats to register\\
      \fmadps &  multiply/add of 16 floats\\
      \gatherps &  gather 16, possibly disconnected floats from memory\\
      \swizzleps &  reorder a channel, and duplicate across channels\\
      \storenrngops &  store 16 floats to memory without register rewrite or reordering\\
      \castpstoi & reinterpret 16 floats as integers\\
      \castitops & reinterpret 16 integers as floats\\
      \permute &  permute channels; do not change individual channels\\
      \intmask &  \\
      \loadunpack &  \\
      \hline
    \end{tabular}
  \end{center}
  %
  \caption{List of MIC vectorial instructions used.}
\label{fig:MIC-vect-instr}
\end{figure*}
In the vector instructions, \ttt{ps} refers to a float (4 bytes),
while \ttt{epi32} refers to a 4-byte integer. Although we do not
discuss double precision in this paper, \ttt{ds} refers to an 8 byte
double precision real number (i.e., \loadds).  Each vector register is
broken up into four channels of 128 bits each. It is possible to
interchange these channels at a cost of a "few" \qes{exact numbers?}
cycles, and it is also possible to execute swizzle operations within a
channel. For example, if the 128 bytes of each channel are labelled as
$ABCD$, the vector instruction \ttt{\_mm512\_swizzle\_ps(v,
  \_MM\_SWIZ\_REG\_AAAA)} replaces each channel by $AAAA$.
